---
title: "Project: Forecasting Product Demand in R"
author: "Manuela da Cruz Chadreque"
date: "02 de Ago de 2020"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: inline
---
# <center>Forecasting demand with time series
### Importing data
ransform the data into an xts object to help with analysis
```{r}
bev<-read.csv("https://www.dropbox.com/s/q5ib3t5ue6tfta0/Bev.csv?dl=1")
head(bev)

# Load xts package 
library(xts)

#Load forecast library
library(forecast) 

# Create the dates object as an index for your xts object
dates <- seq(as.Date("2014-01-19"), length = 176, by = "weeks")

# Create an xts object called bev_xts
bev_xts <- xts(bev, order.by = dates)
head(bev_xts)
```

### Plotting / visualizing data
The column names for the sales of these three products are:

* MET.hi(high end products);
* MET.lo( low end products);
* and MET.sp(specialty products)

```{r}
# Create the individual region sales as their own objects
MET_hi <- bev_xts[,"MET.hi"]
MET_lo <- bev_xts[,"MET.lo"]
MET_sp <- bev_xts[,"MET.sp"]

# Sum the region sales together
MET_t <- MET_hi + MET_lo + MET_sp

# Plot the metropolitan region total sales
plot(MET_t)
```

**Split the data into training and validation sets**

```{r}


MET_t_train <- MET_t[index(MET_t) < "2017-01-01"]
MET_t_valid <- MET_t[index(MET_t) >= "2017-01-01"]

# Use auto.arima() function for metropolitan sales
MET_t_model<-auto.arima(MET_t_train)

# Forecast the first 22 weeks of 2017
forecast_MET_t <- forecast(MET_t_model, h = 22)

# Plot this forecast #
plot(forecast_MET_t)


# Convert to numeric for ease
for_MET_t <- as.numeric(forecast_MET_t$mean)
v_MET_t <- as.numeric(MET_t_valid)

# Calculate the MAE
MAE <- mean(abs(for_MET_t - v_MET_t))

# Calculate the MAPE
MAPE <- 100*mean(abs(for_MET_t - v_MET_t)/v_MET_t)

# Print to see how good your forecast is!
print(MAE)
print(MAPE)

# Convert your forecast to an xts object
for_dates <- seq(as.Date("2017-01-01"), length = 22, by = "weeks")
for_MET_t_xts <- xts(forecast_MET_t$mean, order.by = for_dates)

# Plot the validation data set
plot(MET_t_valid, main = 'Forecast Comparison', ylim = c(4000, 8500))

# Overlay the forecast of 2017
lines(for_MET_t_xts, col = "blue")

# No wonder our MAPE was over 17% on average!

# Confidence Intervals for Forecast
# Plot the validation data set
plot(MET_t_valid, main = 'Forecast Comparison', ylim = c(4000, 8500))

# Overlay the forecast of 2017
lines(for_MET_t_xts, col = "blue")

# Convert the limits to xts objects
lower <- xts(forecast_MET_t$lower[,2], order.by = for_dates)
upper <- xts(forecast_MET_t$upper[,2], order.by = for_dates)

# Adding confidence intervals of forecast to plot
lines(lower, col = "blue", lty = "dashed")
lines(upper, col = "blue", lty = "dashed")

```
 
Although inside the confidence intervals, I bet there are other things that we can include in our model to make our forecasts better!

 

## Calculating price elasticity

```{r}

bev_xts_train <- bev_xts[index(MET_t) <= "2016-12-25"]  
MET_hi<-MET_hi[index(MET_t) <= "2016-12-25"]

# Save the prices of each product
l_MET_hi_p <- as.vector(log(bev_xts_train[,"MET.hi.p"]))

# Save as a data frame
MET_hi_train <- data.frame(as.vector(log(MET_hi)), l_MET_hi_p)
colnames(MET_hi_train) <- c("log_sales", "log_price")

# Calculate the regression
model_MET_hi <- lm(log_sales ~ log_price, data = MET_hi_train)
model_MET_hi
```

Looks like our products are rather elastic here. We better be careful if we increase prices too much!

### Seasonal / holiday / promotional effects
```{r}
MET_hi_p<- bev_xts[,"MET.hi.p"]
# Plot the product's sales
plot(MET_hi)

# Plot the product's price
plot(MET_hi_p)
# Create date indices for New Year's week
n.dates <- as.Date(c("2014-12-28", "2015-12-27", "2016-12-25"))

# Create xts objects for New Year's
newyear <- as.xts(rep(1, 3), order.by = n.dates)

# Create date indices for valentines's week
n1.dates <- as.Date(c("2014-02-09", "2015-02-08", "2016-02-07"))

# Create xts objects for valentine's week
valentine <- as.xts(rep(1, 3), order.by = n1.dates)

# Create date indices for christmas's week
n2.dates <- as.Date(c("2014-12-21", "2015-12-20", "2016-12-18"))

# Create xts objects for christmas's
christmas <- as.xts(rep(1, 3), order.by = n2.dates)

# Create date indices for New mother's week
n3.dates <- as.Date(c("2014-05-04", "2015-05-03", "2016-05-01"))

# Create xts objects for mother's
mother <- as.xts(rep(1, 3), order.by = n3.dates)

# Create sequence of dates for merging
dates_train <- seq(as.Date("2014-01-19"), length = 154, by = "weeks")

# Merge training dates into New Year's object
newyear <- merge(newyear, dates_train, fill = 0)
valentine <- merge(valentine, dates_train, fill = 0)
christmas<- merge(christmas, dates_train, fill = 0)
mother<- merge(mother, dates_train, fill = 0)
# Create MET_hi_train_2 by adding newyear
MET_hi_train_2 <- data.frame(MET_hi_train, as.vector(newyear))
MET_hi_train_2 <- data.frame(MET_hi_train_2, as.vector(valentine))
MET_hi_train_2 <- data.frame(MET_hi_train_2, as.vector(christmas))
MET_hi_train_2 <- data.frame(MET_hi_train_2, as.vector(mother))

colnames(MET_hi_train_2)[3:6] <- c("newyear","valentine","christmas","mother")
head(MET_hi_train_2)
# Build regressions for the product
model_MET_hi_full <- lm(log_sales ~log_price + newyear+valentine+christmas+mother, data = MET_hi_train_2)
summary(model_MET_hi_full)
```
 
For a level of significance of 5%, we can say that the Valentine's Day and  Mother's Day promotion the doesn't lead to a significant increase in sales.

## Forecasting with regression

```{r}
bev_xts_valid<-bev_xts[index(MET_t) >="2017-01-01"]

# Subset the validation prices
l_MET_hi_p_valid <- as.vector(log(bev_xts_valid[,"MET.hi.p"]))

# Create a validation data frame
MET_hi_valid <- data.frame(l_MET_hi_p_valid)
colnames(MET_hi_valid) <- "log_price"

model_MET_hi<-lm(formula = log_sales ~ log_price, data = MET_hi_train)
# Predict the log of sales for your high end product
pred_MET_hi <- predict(model_MET_hi, MET_hi_valid)

# Convert predictions out of log scale
pred_MET_hi <- exp(pred_MET_hi)

# Convert to an xts object
dates_valid <- seq(as.Date("2017-01-01"), length = 22, by = "weeks")
pred_MET_hi_xts <- xts(pred_MET_hi, order.by = dates_valid)

# Plot the forecast
plot(pred_MET_hi_xts)

# Calculate and print the MAPE
MET_hi_v <- bev_xts_valid[,"MET.hi"]

MAPE <- 100*mean(abs((pred_MET_hi_xts - MET_hi_v)/MET_hi_v))
print(MAPE)

```
### Calculating residuals from regression

```{r}
# Calculate the residuals from the model
MET_hi_full_res <- residuals(model_MET_hi_full)

# Convert the residuals to an xts object
MET_hi_full_res <- xts(MET_hi_full_res, order.by = dates_train)
# Plot the histogram of the residuals
par(mfrow=c(2,1))
hist(MET_hi_full_res)

# Plot the residuals over time
plot(MET_hi_full_res)

#Forecasting residuals
#Build an ARIMA model on the residuals: MET_hi_arima
MET_hi_arima<-auto.arima(MET_hi_full_res)

# Look at a summary of the model
summary(MET_hi_arima)

# Forecast 22 weeks with your model: for_MET_hi_arima
for_MET_hi_arima<-forecast(MET_hi_arima,h=22)

# Print first 10 observations
head(for_MET_hi_arima, n = 10)

# Convert your forecasts into an xts object
dates_valid <- seq(as.Date("2017-01-01"), length = 22, by = "weeks")
for_MET_hi_arima <- xts(for_MET_hi_arima$mean , order.by = dates_valid)

# Plot the forecast
plot(for_MET_hi_arima)

# Transfer functions & ensembling

# Convert your residual forecast to the exponential version
for_MET_hi_arima <- exp(for_MET_hi_arima)

# Multiply your forecasts together!
for_MET_hi_final <- pred_MET_hi_xts * for_MET_hi_arima

# Plot the final forecast - don't touch the options!
plot(for_MET_hi_final, ylim = c(1000, 4300))

# Overlay the validation data set
lines(MET_hi_v, col = "blue")

# Calculate the MAE
MAE <- mean(abs(for_MET_hi_final - MET_hi_v))
print(MAE)

# Calculate the MAPE
MAPE <- 100*mean(abs((for_MET_hi_final - MET_hi_v)/MET_hi_v))
print(MAPE)


# Build an ARIMA model using the auto.arima function
MET_hi_model_arima <- auto.arima(MET_hi)

# Forecast the ARIMA model
for_MET_hi <- forecast(MET_hi_model_arima, h = 22)
## Save the forecast as an xts object
dates_valid <- seq(as.Date("2017-01-01"), length = 22, by = "weeks")
for_MET_hi_xts <- xts(for_MET_hi$mean, order.by = dates_valid)

# Calculate the MAPE of the forecast
MAPE <- 100*mean(abs((for_MET_hi_xts-MET_hi_v)/MET_hi_v))
print(MAPE)

# Ensemble the two forecasts together
for_MET_hi_en <- 0.5*(for_MET_hi_xts+pred_MET_hi_xts)

# Calculate the MAE and MAPE
MAE <- mean(abs(for_MET_hi_en-MET_hi_v))
print(MAE)

MAPE <-100* mean(abs((for_MET_hi_en-MET_hi_v)/MET_hi_v))
print(MAPE)

#that didn't improve the MAPE as much as we would hope
```

## Bottom-Up Hierarchical Forecasting

```{r}
# Build a time series model
MET_sp_model_arima <- auto.arima(MET_sp)

# Forecast the time series model for 22 periods
for_MET_sp <- forecast(MET_sp_model_arima, h = 22)

# Create an xts object
for_MET_sp_xts <- xts(for_MET_sp$mean, order.by = dates_valid)

# Calculate the MAPE

mape<-function(yhat, y)
   {
  mean(abs((y - yhat)/y))*100
}

MET_sp_v<-bev_xts_valid[,"MET.sp"]
MAPE <- mape(for_MET_sp_xts, MET_sp_v)
print(MAPE)

MET_sp_train<-xts(MET_hi_train_2, order.by =dates_train )
head(MET_sp_train)


# Build a regression model
model_MET_sp_full <- lm(log_sales~log_price+christmas+valentine+newyear+mother, data=MET_sp_train)

# Create sequence of dates for merging
dates_train <- seq(as.Date("2014-01-19"), length = 154, by = "weeks")
dates_valid <- seq(as.Date("2017-01-01"), length = 22, by = "weeks")

# Subset the validation prices
l_MET_sp_p_valid <- as.vector(log(bev_xts_valid[,"MET.sp.p"]))
l_MET_sp_p_valid
# Create a validation data frame
MET_sp_valid <- data.frame(l_MET_sp_p_valid)
colnames(MET_sp_valid) <- "log_price"

# Merge training dates into New Year's object
newyear <- merge(newyear, dates_valid, fill = 0)
valentine <- merge(valentine, dates_valid, fill = 0)
christmas<- merge(christmas, dates_valid, fill = 0)
mother<- merge(mother, dates_valid, fill = 0)
# Create MET_hi_train_2 by adding newyear
MET_sp_valid <- data.frame(MET_sp_valid, as.vector(newyear))
MET_sp_valid <- data.frame(MET_sp_valid, as.vector(valentine))
MET_sp_valid <- data.frame(MET_sp_valid, as.vector(christmas))
MET_sp_valid <- data.frame(MET_sp_valid, as.vector(mother))

colnames(MET_sp_valid)[2:5] <- c("newyear","valentine","christmas","mother")
head(MET_sp_valid)


# Forecast the regression model
pred_MET_sp <- forecast(model_MET_sp_full,MET_sp_valid )
head(MET_sp_valid)
# Exponentiate your predictions and create an xts object
pred_MET_sp <- exp(pred_MET_sp$mean)
length(pred_MET_sp)
dates_valid
head(pred_MET_sp,100)
# pred_MET_sp_xts <- xts(pred_MET_sp, order.by=dates_valid)
# 
# # # Calculate MAPE
#  MAPE <- mape(pred_MET_sp_xts, MET_sp_v)
#  print(MAPE)
# # 
#  # Ensemble the two forecasts
#  for_MET_sp_en <- 0.5*(for_MET_sp_xts+pred_MET_sp_xts)
# # 
# # # Calculate the MAPE
#  MAPE <- mape(for_MET_sp_en, MET_sp_v)
#  print(MAPE)
```


