---
title: "Time Series Analysis Part 1"
author: "Manuela da Cruz Chadreque"
date: "25 de Abril de 2019"
output:
  html_document:
    keep_md: yes
    number_sections: yes
    theme: journal
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r eval=FALSE, include=FALSE}
setwd("C:/Users/maiam/Dropbox/PROFESSIONAL DEVELOPMENT/DATA SCIENCE/01_R/TIME SERIES ANALYSIS/Introduction to Time Series Analysis")
```

## Time Series  Exploratory Analysis


##simple Forecasting Methods
###Average Method
The forecast of all future values are equal to the mean 
$$ \hat{x}_{n+k|n}=\bar{x}=\frac{x_1+x_2+...+x_n}{n} $$

###Naive method
Forecasts equal to last observed value 
 $$ \hat{x}_{n+k|n}=x_n $$
###Seasonal naive method
Forecasts equal to last value from same season. 
$$ \hat{x}_{n+k|n}=x_{n+k-lm} $$
where: 
m is the seasonal period
$$I= \frac{(k-1)}{m} +1$$
 Drift method 
Forecasts equal to last value plus average change. Equivalent extrapolating a line drawn between first and last observations.

$$ \hat{x}_{n+k|n}=x_{n}+ \frac{(k-1)}{m} \sum_{i=2}^n{x_t-x_{t-1}}=x_n+\frac{k}{n-1}(x_n-x_1) $$
##Decomposition methods
Decomposition methods are based in the decomposition of three movements or strengths: **trend**, **seasonality** and **noise**:  

$$X_t=f(T_t,S_t,e_t)$$
##ARIMA models (Autoregressive Integrated Moving Average) 
Models that relate the present value of a series to past values and past prediction errors

###Temperatures
```{r}
plot.ts(window(x=ts.temp),type="b",xlab="Time",ylab="Temperatures", xlim=c(2014,2018))
lines(meanf(window(ts.temp),h=24)$mean,col=2,lwd=3)
lines(naive(window(ts.temp),h=24)$mean,col=3,lwd=3)
lines(snaive(window(ts.temp),h=24)$mean,col=4,lwd=3)
lines(rwf(window(ts.temp),drift=T,h=24)$mean,col=5,lwd=3)
text<-c("Mean method","Naive method", "Seasonal naive method", "Drift method")
legend("topright",text,col=2:5,lty=1,ncol=1,cex=1,lwd=3)
```
 
analysis

```{r}
plot.ts(window(bricksq),type="b",xlab="Time",ylab="Quarterly production of bricks (in millions of units) at Portland, Australia") 
lines(meanf(window(bricksq),h=24)$mean,col=2,lwd=3)
lines(naive(window(bricksq),h=24)$mean,col=3,lwd=3)
lines(snaive(window(bricksq),h=24)$mean,col=4,lwd=3)
lines(rwf(window(bricksq),drift=T,h=24)$mean,col=5,lwd=3)
text<-c("Mean method","Naive method", "Seasonal naive method", "Drift method")
legend("topright",text,col=2:5,lty=1,ncol=1,cex=1,lwd=3)
```
 
analysis


###Annual Electric Sales for South Australia in GWh
```{r}
plot.ts(window(elecsales),type="b",xlab="Time",ylab="Annual Electric Sales for South Australia in GWh")
lines(meanf(window(elecsales),h=24)$mean,col=2,lwd=3)
lines(naive(window(elecsales),h=24)$mean,col=3,lwd=3)
lines(snaive(window(elecsales),h=24)$mean,col=4,lwd=3)
lines(rwf(window(elecsales),drift=T,h=24)$mean,col=5,lwd=3)
text<-c("Mean method","Naive method", "Seasonal naive method", "Drift method")
legend("topright",text,col=2:5,lty=1,ncol=1,cex=1,lwd=3)
```
 
analysis



###Monthly sales of new one-family houses sold in the USA

```{r}
plot.ts(window(hsales, start=1973, end=1995),type="b",xlab="Time",ylab="Monthly sales of new one-family houses sold in the USA") #time series plot of temp with points market as bubbles

lines(meanf(window(hsales),h=24)$mean,col=2,lwd=3)
lines(naive(window(hsales),h=24)$mean,col=3,lwd=3)
lines(snaive(window(hsales),h=24)$mean,col=4,lwd=3)
lines(rwf(window(hsales),drift=T,h=24)$mean,col=5,lwd=3)
text<-c("Mean method","Naive method", "Seasonal naive method", "Drift method")
legend("topright",text,col=2:5,lty=1,ncol=1,cex=1,lwd=3)
```
 
analysis

 Daily closing IBM stock prices
```{r}
plot.ts(window(ibmclose),type="b",xlab="Time",ylab="Daily closing IBM stock prices", xlim=c(1, 379)) #time series plot of temp with points market as bubbles

lines(meanf(window(ibmclose),h=24)$mean,col=2,lwd=3)
lines(naive(window(ibmclose),h=24)$mean,col=3,lwd=3)
lines(snaive(window(ibmclose),h=24)$mean,col=4,lwd=3)
lines(rwf(window(ibmclose),drift=T,h=24)$mean,col=5,lwd=3)
text<-c("Mean method","Naive method", "Seasonal naive method", "Drift method")
legend("topright",text,col=2:5,lty=1,ncol=1,cex=1,lwd=3)

```
 
References


@Manual{R-base,
  title = {Using R for Time Series Analysis},
  author = {{Avril Coghlan}},
  url = {https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html#time-series-analysis},
}

@Manual{R-base,
  title = {Forecasting: Principles and Practice},
  author = {{Rob J Hyndman, George Athanasopoulos}},
  organization = {Monash University},
  address = {Austria},
  year = {2016},
  url = {https://otexts.com/fpp2/},
}

