---
title: "Tutorial: Introduction to Time Series Analysis"
author: "Manuela da Cruz Chadreque"
date: "12 de Março de 2019"
output: 
  html_document: 
    theme: cerulean
---


```{r}
setwd("C:/Users/maiam/Dropbox/PROFESSIONAL DEVELOPMENT/DATA SCIENCE/01_R/TIME SERIES ANALYSIS/Introduction to Time Series Analysis")
```
# Exploratory time series data analysis


```{r}
# Print the Nile dataset
print(Nile)

# List the number of observations in the Nile dataset
length(Nile)

# Display the first 10 elements of the Nile dataset
head(Nile, n=10)

# Display the last 12 elements of the Nile dataset
tail(Nile,n=12)

```
Basic time series plots

```{r}
# Plot the Nile data
plot(Nile)

# Plot the Nile data with xlab and ylab arguments
plot(Nile, xlab = "Year", ylab = "River Volume (1e9 m^{3})")

# Plot the Nile data with xlab, ylab, main, and type arguments
plot(Nile, xlab = "Year", ylab = "River Volume (1e9 m^{3})",main = "Annual River Nile Volume at Aswan, 1871-1970", type ="b")
```

**Ploting Air Passagers 

```{r}
# Plot AirPassengers
plot(AirPassengers)

# View the start and end dates of AirPassengers
start(AirPassengers)
end(AirPassengers)

# Use time(), deltat(), frequency(), and cycle() with AirPassengers 

time(AirPassengers)
deltat(AirPassengers)
frequency(AirPassengers)
cycle(AirPassengers)

# Plot the AirPassengers data
plot(AirPassengers)

# Compute the mean of AirPassengers

mean(AirPassengers)
# Impute mean values to NA in AirPassengers
AirPassengers[85:96] <- mean(AirPassengers, na.rm =TRUE )

# Generate another plot of AirPassengers
plot(AirPassengers)

# Add the complete AirPassengers data to your plot
rm(AirPassengers)
points(AirPassengers, type = "l", col = 2, lty = 3)

```

####Creating a time series object with ts()

```{r}
data_vector<-read.csv("C:/Users/maiam/Dropbox/PROFESSIONAL DEVELOPMENT/DATA SCIENCE/01_R/TIME SERIES ANALYSIS/Introduction to Time Series Analysis/data.csv")
# Use print() and plot() to view data_vector
head(data_vector)
plot(data_vector)

# Convert data_vector to a ts object with start = 2004 and frequency = 4
time_series <- ts(data_vector,start=2004,frequency=4)

# Use print() and plot() to view time_series
print(time_series)
plot(time_series)

```
####Testing whether an object is a time series

```{r}
# Check whether data_vector and time_series are ts objects
is.ts(time_series)
is.ts(data_vector)

# Check whether Nile is a ts object
is.ts(Nile)

# Check whether AirPassengers is a ts object

is.ts(AirPassengers)

```

####Plotting a time series object

```{r}
# Check whether eu_stocks is a ts object
eu_stocks<-EuStockMarkets
is.ts(eu_stocks)

# View the start, end, and frequency of eu_stocks
start(eu_stocks)
end(eu_stocks)
frequency(eu_stocks)

# Generate a simple plot of eu_stocks
plot(eu_stocks)

# Use ts.plot with eu_stocks
ts.plot(eu_stocks, col = 1:4, xlab = "Year", ylab = "Index Value", main = "Major European Stock Indices, 1991-1998")

# Add a legend to your ts.plot
legend("topleft", colnames(eu_stocks), lty = 1, col = 1:4, bty = "n")

```
## Trend Spotting!
### Removing trends in variability via the logarithmic transformation

```{r}
rapid_growth<-read.csv("rapid_growth.csv")
rapid_growth<-ts(rapid_growth)
is.ts(rapid_growth)
# Log rapid_growth
linear_growth <-log(rapid_growth)
  
# Plot linear_growth using ts.plot()
 ts.plot(linear_growth)
 
```
##Removing trends in level by differencing

```{r}
z<-read.csv("z.csv")
z<-ts(z)
# Generate the first difference of z
dz <- diff(z)
  length(z)
# Plot dz
ts.plot(dz)

# View the length of z and dz, respectively
length(z)
length(dz)

```
##Removing seasonal trends with seasonal differencing

```{r}
x<-read.csv("x.csv")
# # Generate a diff of x with lag = 4. Save this to dx
dx<-ts(x)
dx <- diff(dx,lag=4)
length(dx)
# # Plot dx
ts.plot(dx)

# # View the length of x and dx, respectively 
length(x)
length(dx)

```
Simulate the white noise model

```{r}
# Simulate a WN model with list(order = c(0, 0, 0))
white_noise <- arima.sim(model = list(order = c(0, 0, 0)), n = 100)

# Plot your white_noise data
ts.plot(white_noise)


# Simulate from the WN model with: mean = 100, sd = 10
white_noise_2 <- arima.sim(model = list(order = c(0, 0, 0)), n = 100, mean =100, sd = 10)

# Plot your white_noise_2 data
ts.plot(white_noise_2)



```

Estimate the white noise model


```{r}
y<-read.csv("y.csv")
y<-ts(y)

# # Fit the WN model to y using the arima command
arima(y,order=c(0,0,0))
# # Calculate the sample mean and sample variance of y
mean(y)
var(y)


```
##3. Correlation analysis and the autocorrelation function


```{r}
# Plot eu_stocks
plot(eu_stocks)

# Use this code to convert prices to returns
returns <- eu_stocks[-1,] / eu_stocks[-1860,] - 1

# Convert returns to ts
returns <- ts(returns, start = c(1991, 130), frequency =260)

# Plot returns

plot(returns)
# Use this code to convert prices to log returns
logreturns <- diff(log(eu_stocks))

# Plot logreturns
plot(logreturns)

```
Characteristics of financial time series


```{r}
# Generate means from eu_percentreturns
eu_percentreturns<-returns*100
colMeans(eu_percentreturns) 

# Use apply to calculate sample variance from eu_percentreturns
apply(eu_percentreturns, MARGIN = 2, FUN = var)

# Use apply to calculate standard deviation from eu_percentreturns
apply(eu_percentreturns, MARGIN = 2, FUN = sd)

# Display a histogram of percent returns for each index
par(mfrow = c(2,2))
apply(eu_percentreturns, MARGIN = 2, FUN = hist, main = "", xlab = "Percentage Return")

# Display normal quantile plots of percent returns for each index
par(mfrow = c(2,2))
apply(eu_percentreturns, MARGIN = 2, FUN = qqnorm, main = "")
qqline(eu_percentreturns)

```

Plotting pairs of data

```{r}
DAX<-eu_stocks[,1]
FTSE<-eu_stocks[,3] 
# Make a scatterplot of DAX and FTSE
plot(DAX, FTSE)

# Make a scatterplot matrix of eu_stocks
pairs(eu_stocks)

# Convert eu_stocks to log returns
logreturns <- diff(log(eu_stocks))

# Plot logreturns
plot(logreturns)

# Make a scatterplot matrix of logreturns
pairs(logreturns)

```
Calculating sample covariances and correlations

```{r}
x<-read.csv("x1.csv")
x<-ts(x)
n<-length(x)
n
# Define x_t0 as x[-1]
x_t0 <- x[-1]

# Define x_t1 as x[-n]
x_t1 <- x[-n]

# Confirm that x_t0 and x_t1 are (x[t], x[t-1]) pairs  
head(cbind(x_t0, x_t1))

# Plot x_t0 and x_t1
plot(x_t0,x_t1)

# View the correlation between x_t0 and x_t1
cor(x_t0,x_t1)

# Use acf with x
acf(x, lag.max = 1, plot =FALSE)

# Confirm that difference factor is (n-1)/n
cor(x_t1, x_t0) * (n-1)/n

```

The autocorrelation function


```{r}

# Generate ACF estimates for x up to lag-10
acf(x, lag.max =10, plot = FALSE)

# Type the ACF estimate at lag-10 
acf(x, lag.max =0.100, plot = FALSE)


# Type the ACF estimate at lag-5
acf(x, lag.max =0.198, plot = FALSE)

```

Visualizing the autocorrelation function


```{r}
# View the ACF of x
acf(x)

# View the ACF of y
acf(y)

# View the ACF of z
acf(z)

```
##4. Autoregression
Simulate the autoregressive model

```{r}
# Simulate an AR model with 0.5 slope
x <- arima.sim(model = list(ar=0.5), n = 100)

# Simulate an AR model with 0.9 slope
y <- arima.sim(model=list(ar=0.9),n=100)

# Simulate an AR model with -0.75 slope
z <- arima.sim(model=list(ar=-0.75),n=100)

# Plot your simulated data
plot.ts(cbind(x,y, z))

```
Estimate the autocorrelation function (ACF) 


```{r}
# Calculate the ACF for x
acf(x)

# Calculate the ACF for y
acf(y)

# Calculate the ACF for z
acf(z)

```

Compare the random walk (RW) and autoregressive (AR) models

```{r}
# Simulate and plot AR model with slope 0.9 
x <- arima.sim(model = list(ar=0.9), n = 200)
ts.plot(x)
acf(x)

# Simulate and plot AR model with slope 0.98
y <- arima.sim(model=list(ar=0.98),n=200)
ts.plot(y)
acf(y)


# Simulate and plot RW model
z <-arima.sim(model = list(order = c(0, 1, 0)), n = 200)
ts.plot(z)
acf(z)

```
Estimate the autoregressive (AR) model

```{r}
# Fit the AR model to x
arima(x, order =c(1,0,0))

# Copy and paste the slope (ar1) estimate
0.8575

# Copy and paste the slope mean (intercept) estimate
-0.0948

# Copy and paste the innovation variance (sigma^2) estimate
1.022
# Fit the AR model to AirPassengers
AR <-arima(AirPassengers,order=c(1,0,0))
print(AR)

# Run the following commands to plot the series and fitted values
ts.plot(AirPassengers)
AR_fitted <- AirPassengers - residuals(AR)
points(AR_fitted, type = "l", col = 2, lty = 2)


```
Simple forecasts from an estimated AR model

```{r}
# Simulate an AR model with 0.5 slope
x <- arima.sim(model = list(ar=0.5), n = 100)

# Simulate an AR model with 0.9 slope
y <- arima.sim(model=list(ar=0.9),n=100)

# Simulate an AR model with -0.75 slope
z <- arima.sim(model=list(ar=-0.75),n=100)

# Plot your simulated data
plot.ts(cbind(x,y, z))


# Simulate and plot AR model with slope 0.9 
x <- arima.sim(model = list(ar=0.9), n = 200)
ts.plot(x)
acf(x)

# Simulate and plot AR model with slope 0.98
y <- arima.sim(model=list(ar=0.98),n=200)
ts.plot(y)
acf(y)


# Simulate and plot RW model
z <-arima.sim(model = list(order = c(0, 1, 0)), n = 200)
ts.plot(z)
acf(z)

# Fit an AR model to Nile
AR_fit <-arima(Nile, order  =c(1,0,0))
print(AR_fit)

# Use predict() to make a 1-step forecast
predict_AR<- predict(AR_fit, n.ahead = 1)

# Obtain the 1-step forecast using $pred[1]
predict_AR$pred[1]


# Use predict to make 1-step through 10-step forecasts
predict(AR_fit, n.ahead = 10)

# Fit an AR model to Nile
AR_fit <-arima(Nile, order  =c(1,0,0))
print(AR_fit)

# Use predict() to make a 1-step forecast
predict_AR<- predict(AR_fit, n.ahead = 1)

# Obtain the 1-step forecast using $pred[1]
predict_AR$pred[1]

```
  
##5. A simple moving average

```{r}
# Generate MA model with slope 0.5
x <- arima.sim(model = list(ma=0.5), n =100)

# Generate MA model with slope 0.9
y <- arima.sim(model=list(ma=0.9),n=100)

# Generate MA model with slope -0.5
z <- arima.sim(model=list(ma=-0.5),n=100)

# Plot all three models together
plot.ts(cbind(x,y,z))

```
Estimate the autocorrelation function (ACF) for a moving average


```{r}
# Calculate ACF for x
acf(x)

# Calculate ACF for y
acf(y)

# Calculate ACF for z
acf(z)


```
Estimate the simple moving average model

```{r}
# Fit the MA model to x
arima(x, order = c(0,0,1))

# Paste the slope (ma1) estimate below
0.7928

# Paste the slope mean (intercept) estimate below
0.1589

# Paste the innovation variance (sigma^2) estimate below
0.9576

# Fit the MA model to Nile
MA <- arima(Nile, order =c(0,0,1))
print(MA)

# Plot Nile and MA_fit 
ts.plot(Nile)
MA_fit <- Nile - resid(MA)
points(MA_fit, type = "l", col = 2, lty = 2)

```
Simple forecasts from an estimated MA model

```{r}
# Make a 1-step forecast based on MA
predict_MA <-predict(MA,n.ahead=1)

# Obtain the 1-step forecast using $pred[1]
predict_MA$pred[1]

# Make a 1-step through 10-step forecast based on MA
predict(MA,n.ahead=10)

# Plot the Nile series plus the forecast and 95% prediction intervals
ts.plot(Nile, xlim = c(1871, 1980))
MA_forecasts <- predict(MA, n.ahead = 10)$pred
MA_forecast_se <- predict(MA, n.ahead = 10)$se
points(MA_forecasts, type = "l", col = 2)
points(MA_forecasts - 2*MA_forecast_se, type = "l", col = 2, lty = 2)
points(MA_forecasts + 2*MA_forecast_se, type = "l", col = 2, lty = 2)

```
AR vs MA models


```{r}
AR= arima(x = Nile, order = c(1, 0, 0))
MA=arima(x = Nile, order = c(0, 0, 1))
# # Find AIC of AR
AIC(AR)
# 
# # Find AIC of MA
AIC(MA)
# 
# # Find BIC of AR
BIC(AR)
# 
# # Find BIC of MA
BIC(MA)

```

