---
title: "Quantitative Analyst with R"
author: "Manuela Chadreque"
date: "11 de MarÃfÂ§o de 2019"
output:
  html_document: default
  pdf_document: default
---

#instal and load packedges


## Introduction to R for Finance!

###The Basics
####Arithmetic in R
Addition: +
Subtraction: -
Multiplication: *
Division: /
Exponentiation: ^ or **
Modulo: %%
```{r}
# Addition 
2 + 2

# Subtraction
4 - 1

# Multiplication
3 * 4

# Division

4/2
# Exponentiation

2^4
# Modulo
7%%3
```

###Assignment and variables
A variable allows you to store a value or an object in R. You can then later use this variable's name to easily access the value or the object that is stored within this variable. You use <- to assign a variable:

my_money <- 100


```{r}
# Assign 200 to savings
savings <- 200

# Print the value of savings to the console
savings
```
####Aritmetrics with variables

```{r}

# Assign 100 to my_money
my_money <- 100

# Assign 200 to dans_money
dans_money<-200

# Add my_money and dans_money
my_money+dans_money

# Add my_money and dans_money again, save the result to our_money
our_money<-my_money+dans_money
print(our_money)
```

#Financial returns 
Assume you have $100. During January, you make a 5% return on that money. How much do you have at the end of January? Well, you have 100% of your starting money, plus another 5%: 100% + 5% = 105%. In decimals, this is 1 + .05 = 1.05. This 1.05 is the return multiplier for January, and you multiply your original $100 by it to get the amount you have at the end of January.

105 = 100 * 1.05

Or in terms of variables:

post_jan_cash <- starting_cash * jan_ret

A quick way to get the multiplier is:

multiplier = 1 + (return / 100)
 If, in February, you earn another 2% on your cash, how would you calculate the total amount at the end of February? You already know that the amount at the end of January is $100 * 1.05 = $105. To get from the end of January to the end of February, just use another multiplier!

$105 * 1.02 = $107.1

Which is equivalent to:

$100 * 1.05 * 1.02 = $107.1

```{r}

# Variables for starting_cash and 5% return during January
starting_cash <- 200
jan_ret <- 5
jan_mult <- 1 + (jan_ret / 100)

# How much money do you have at the end of January?
post_jan_cash <- starting_cash*jan_mult

# Print post_jan_cash
post_jan_cash

# January 10% return multiplier
jan_ret_10 <- 10
jan_mult_10 <- 1+(jan_ret_10/100)

# How much money do you have at the end of January now?
post_jan_cash_10 <- starting_cash*jan_mult_10


# Print post_jan_cash_10
post_jan_cash_10

 
# Starting cash and returns 
starting_cash <- 200
jan_ret <- 4
feb_ret <- 5

# Multipliers
jan_mult <- 1+(jan_ret/100)
feb_mult <- 1+feb_ret/100

# Total cash at the end of the two months
total_cash <- starting_cash*jan_mult*feb_mult

# Print total_cash
total_cash
```
###Data type exploration
To get started, here are some of R's most basic data types:

Numerics are decimal numbers like 4.5. A special type of numeric is an integer, which is a numeric without a decimal piece. Integers must be specified like 4L.
Logicals are the boolean values TRUE and FALSE. Capital letters are important here; true and false are not valid.
Characters are text values like "hello world".

####What's that data type?
Up until now, you have been determining what data type a variable is just by looks. There is actually a better way to check this.

class(my_var)

```{r}
# 
# # company's stock price is a numeric
# company_stock <- 150.45
# 
# # Bond credit ratings are characters
# credit_rating <-"AAA"
# 
# # You like the stock market. TRUE or FALSE?
# my_answer <- TRUE
# 
# # Print my_answer
# my_answer

```

###Vector
####c()ombine
Now is where things get fun! It is time to create your first vector. Since this is a finance oriented course, it is only appropriate that your first vector be a numeric vector of stock prices. Remember, you create a vector using the combine function, c(), and each element you add is separated by a comma.
We can also add names to each return in your vector. You do this using names()
```{r}

# Another numeric vector
ibm_stock <- c(159.82, 160.02, 159.84)

# Another character vector
finance <-c("stocks", "bonds", "investments")

# A logical vector
logic <- c(TRUE,FALSE,TRUE)


# Vectors of 12 months of returns, and month names
ret <- c(5, 2, 3, 7, 8, 3, 5, 9, 1, 4, 6, 3)
months <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

# Add names to ret
names(ret) <- months

# Print out ret to see the new names!
ret

```

###Visualize your vector
The plot() function is one of the many ways to create a graph from your data in R. Passing in a vector will add its values to the y-axis of the graph, and on the x-axis will be an index created from the order that your vector is in.

Inside of plot(), you can change the type of your graph using type =. The default is "p" for points, but you can also change it to "l" for line.

``` {r}

apple_stock <- c(109.49, 109.90, 109.11, 109.95, 111.03, 112.12)
credit_rating <- c("AAA", "AA", "BBB", "BB", "B")
micr<-c(59.20,59.25,60.22,59.95,61.37,61.01,61.97,62.17,62.98,62.68,62.58,62.30
,63.62,63.54,63.54,63.55,63.24,63.28,62.99,62.90,62.14)

# Look at the data
apple_stock

# Plot the data points
plot(apple_stock)

# Plot the data as a line graph
plot(apple_stock, type = "l")

```
####Weighted average
As a finance professional, there are a number of important calculations that you will have to know. One of these is the weighted average. The weighted average allows you to calculate your portfolio return over a time period. Consider the following example:

Assume you have 40% of your cash in Apple stock, and 60% of your cash in IBM stock. If, in January, Apple earned 5% and IBM earned 7%, what was your total portfolio return?

To calculate this, take the return of each stock in your portfolio, and multiply it by the weight of that stock. Then sum up all of the results. For this example, you would do:

6.2 = 5 * .4 + 7 * .6

Or, in variable terms:

portf_ret <- apple_ret * apple_weight + ibm_ret * ibm_weight

```{r}

# Weights and returns
micr_ret<- 7
sony_ret <- 9
micr_weight <- .2
sony_weight <- .8

# Portfolio return
portf_ret <-micr_ret * micr_weight + sony_ret * sony_weight


# Weights, returns, and company names
ret <- c(7, 9)
weight <- c(.2, .8)
companies <- c("Microsoft", "Sony")

# Assign company names to your vectors
names(ret) <- companies
names(weight) <- companies

# Multiply the returns and weights together 
ret_X_weight <- ret*weight

# Print ret_X_weight
ret_X_weight

# Sum to get the total portfolio return
portf_ret <-sum(ret_X_weight)

# Print portf_ret
sum(portf_ret)


# Print ret
ret 

# Assign 1/3 to weight
weight <- 1/3

# Create ret_X_weight
ret_X_weight <- ret*weight

# Calculate your portfolio return
portf_ret <- sum(ret_X_weight)

# Vector of length 3 * Vector of length 2?
ret * c(.2, .6)
```

####Vector subsetting
Sometimes, you will only want to use specific pieces of your vectors, and you'll need some way to access just those parts. For example, what if you only wanted the first month of returns from the vector of 12 months of returns? To solve this, you can subset the vector using [ ].

Here is the 12 month return vector:

ret <- c(5, 2, 3, 7, 8, 3, 5, 9, 1, 4, 6, 3)

Select the first month: ret[1].

Select the first month by name: ret["Jan"].

Select the first three months: ret[1:3] or ret[c(1, 2, 3)].


```{r}
ret <- c(5, 2, 3, 7, 8, 3, 5, 9, 1, 4, 6, 3)

# First 6 months of returns
ret[1:6]

# Just March and May
ret[c("Mar","May")]

# Omit the first month of returns
ret[-1]
```

###Matrix
Matrices are similar to vectors, except they are in 2 dimensions! Let's create a 2x2 matrix "by hand" using matrix().

```{r}

# A vector of 9 numbers
my_vector <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)

# 3x3 matrix
my_matrix <- matrix(data = my_vector, nrow = 3, ncol =3)

# Print my_matrix
my_matrix

# Filling across using byrow = TRUE
apple_micr_matrix<-matrix(data = c(2, 3, 4, 5), nrow = 2, ncol = 2, byrow = TRUE)

# View the data
apple_micr_matrix

# Scatter plot of Microsoft vs Apple
plot(apple_micr_matrix)

```
###cor()relation
The cor() function will calculate the correlation between two vectors, or will create a correlation matrix when given a matrix.

```{r}
apple <- c(109.49, 109.90, 109.11, 109.95, 111.03)
ibm <- c(159.82, 160.02, 159.84, 160.35, 164.79)

# cbind the vectors together
cbind_stocks <- cbind(apple,ibm,micr)

# Print cbind_stocks
cbind_stocks

# rbind the vectors together
rbind_stocks <- rbind(apple,ibm,micr)

# Print rbind_stocks
rbind_stocks

```
###Matrix subsetting
Matrix subsetting
Just like vectors, matrices can be selected from and subsetted! To do this, you will again use [ ], but this time it will have two inputs. The basic structure is:

my_matrix[row, col]

Then: To select the first row and first column of stocks from the last example: stocks[1,1]

To select the entire first row, leave the col empty: stocks[1, ]

To select the first two rows: stocks[1:2, ] or stocks[c(1,2), ]

To select an entire column, leave the row empty: stocks[, 1]

You can also select an entire column by name: stocks[, "apple"]

```{r}
stocks<-cbind_stocks
# Third row
stocks[3,]

# Fourth and fifth row of the ibm column
stocks[c(4,5),"ibm"]

# apple and micr columns
stocks[,c("apple","micr")]

```

###data.frame()
Data frames are great because of their ability to hold a different type of data in each column. To get started, let's use the data.frame() function to create a data frame of your business's future cash flows. Here are the variables that will be in the data frame:

company - The company that is paying you the cash flow (A or B).
cash_flow - The amount of money a company will receive.
year - The number of years from now that you receive the cash flow.

```{r}
# Variables
company <- c("A", "A", "A", "B", "B", "B", "B")
cash_flow <- c(1000, 4000, 550, 1500, 1100, 750, 6000)
year <- c(1, 3, 4, 1, 2, 4, 5)

# Data frame
cash <- data.frame(company,cash_flow,year)

# Print cash
cash

```
###Making head()s and tail()s of your data with some str()ucture
Time to introduce a few simple, but very useful functions.

head() - Returns the first few rows of a data frame. By default, 6. To change this, use head(cash, n = ___)
tail() - Returns the last few rows of a data frame. By default, 6. To change this, use tail(cash, n = ___)
str() - Check the structure of an object. This fantastic function will show you the data type of the object you pass in (here, data.frame), and will list each column variable along with its data type.
With a small data set such as yours, head() and tail() are not incredibly useful, but imagine if you had a data frame of hundreds or thousands of rows!

```{r}
# Call head() for the first 4 rows
head(cash,n=4)

# Call tail() for the last 3 rows
tail(cash,n=3)

# Call str()
str(cash)
```

####Naming your columns / rows
```{r}

# Fix your column names
colnames(cash)<-c("company","cash_flow","year")

# Print out the column names of cash
names(cash)
```
####Accessing and subsetting data frames (1)
Even more often than with vectors, you are going to want to subset your data frame or access certain columns. Again, one of the ways to do this is to use [ ]. The notation is just like matrices! Here are some examples:

Select the first row: cash[1, ]

Select the first column: cash[ ,1]

Select the first column by name: cash[ ,"company"]

Accessing and subsetting data frames (2)
As you might imagine, selecting a specific column from a data frame is a common manipulation. So common, in fact, that it was given its own shortcut, the $. The following return the same answer:

cash$cash_flow

[1] 1000 4000  550 1500 1100  750 6000

cash[,"cash_flow"]

[1] 1000 4000  550 1500 1100  750 6000
Accessing and subsetting data frames (3)
Often, just simply selecting a column from a data frame is not all you want to do. What if you are only interested in the cash flows from company A? For more flexibility, try subset()!

subset(cash, company == "A")

  company cash_flow year
1       A      1000    1
2       A      4000    3
3       A       550    4

```{r}

# Third row, second column
cash[3,2]


# Fifth row of the "year" column
cash[5,"year"]


# Select the year column
cash$year

# Select the cash_flow column and multiply by 2
cash$cash_flow*2

# Delete the company column
cash$company <- NULL

# Print cash again
cash

# Rows about company B
subset(cash,company=="B")

# Rows with cash flows due in 1 year
subset(cash,year==1)

```

###Adding new columns
In a perfect world, you could be 100% certain that you will receive all of your cash flows. But, since these are predictions about the future, there is always a chance that someone won't be able to pay! You decide to run some analysis about a worst case scenario where you only receive half of your expected cash flow. To save the worst case scenario for later analysis, you decide to add it as a new column to the data frame!

cash$half_cash <- cash$cash_flow * .5

cash

  company cash_flow year half_cash
1       A      1000    1       500
2       A      4000    3      2000
3       A       550    4       275
4       B      1500    1       750
5       B      1100    2       550
6       B       750    4       375
7       B      6000    5      3000

```{r}

# Quarter cash flow scenario
cash$quarter_cash <-cash$cash_flow*0.25

# Double year scenario
cash$double_year<-cash$year*2


```
###Present value of projected cash flows
If you expect a cash flow of $100 to be received 1 year from now, what is the present value of that cash flow at a 5% interest rate? To calculate this, you discount the cash flow to get it in terms of today's dollars. The general formula for this is:

present_value <- cash_flow * (1 + interest / 100) ^ -year

95.238 = 100 * (1.05) ^ -1

Another way to think about this is to reverse the problem. If you have $95.238 today, and it earns 5% over the next year, how much money do you have at the end of the year? We know how to do this problem from way back in chapter 1! Find the multiplier that corresponds to 5% and multiply by $95.238!

100 = 95.238 * (1.05)

```{r}

# Present value of $4000, in 3 years, at 5%
present_value_4k <- 4000*(1+0.05)^-3

# Present value of all cash flows
cash$present_value <-cash$cash_flow*(1+0.05)^-cash$year

# Print out cash
cash


# Total present value of cash
total_pv <- sum(cash$present_value)

# Company B information
cash_B <- subset(cash,company=="B")

# Total present value of cash_B
total_pv_B <- sum(cash_B$present_value)

```

###Factor
Bond credit ratings are common in the fixed income side of the finance world as a simple measure of how "risky" a certain bond might be. Here, riskiness can be defined as the probability of default, which means an inability to pay back your debts. The Standard and Poor's and Fitch credit rating agency has defined the following ratings, from least likely to default to most likely:

AAA, AA, A, BBB, BB, B, CCC, CC, C, D
This is a perfect example of a factor! It is a categorical variable that takes on a limited number of levels.

To create a factor in R, use the factor() function, and pass in a vector that you want to be converted into a factor.

Suppose you have a portfolio of 7 bonds with these credit ratings:

credit_rating <- c("AAA", "AA", "A", "BBB", "AA", "BBB", "A")

To create a factor from this:

factor(credit_rating)

[1] AAA AA  A   BBB AA  BBB A  
Levels: A AA AAA BBB

```{r}

# credit_rating character vector
credit_rating <- c("BB", "AAA", "AA", "CCC", "AA", "AAA", "B", "BB")

# Create a factor from credit_rating
credit_factor <- factor(credit_rating)

# Print out your new factor
credit_factor

# Call str() on credit_rating
str(credit_rating)

# Call str() on credit_factor
str(credit_factor)


# Identify unique levels
levels(credit_factor)

# Rename the levels of credit_factor
levels(credit_factor)<-c("2A","3A","1B","2B","3C")

# Print credit_factor
credit_factor
```

###Factor summary

As any good bond investor would do, you would like to keep track of how many bonds you are holding of each credit rating. A way to present a table of the counts of each bond credit rating would be great! Luckily for you, the summary() function for factors can help you with that.


```{r}

# Summarize the character vector, credit_rating
summary(credit_rating)

# Summarize the factor, credit_factor
summary(credit_factor)
```

###Visualize your factor
You can also visualize the table that you created in the last example by using a bar chart. A bar chart is a type of graph that displays groups of data using rectangular bars where the height of each bar represents the number of counts in that group.

The plot() function can again take care of all of the magic for you, check it out!

```{r}

# Visualize your factor!
plot(credit_factor)

```
####Bucketing a numeric variable into a factor
Your old friend Dan sent you a list of 50 AAA rated bonds called AAA_rank, with each bond having an additional number from 1-100 describing how profitable he thinks that bond will be (100 being the most profitable). You are interested in doing further analysis on his suggestions, but first it would be nice if the bonds were bucketed by their ranking somehow. This would help you create groups of bonds, from least profitable to most profitable, to more easily analyze them.

This is a great example of creating a factor from a numeric vector. The easiest way to do this is to use cut(). Below, Dan's 1-100 ranking is bucketed into 5 evenly spaced groups. Note that the ( in the factor levels means we do not include the number beside it in that group, and the ] means that we do include that number in the group.

```{r}
AAA_rank<-c(31,48,100,53,85,73,62,74,42,38,97,61,48,86,44,9,43,18,62,31,48,100,53,85,73,62,74,42,38,97,61,48,86,44,9,43,18,62,31,48,100,53,85,73,62,74,42,38,97,61,48,86,44,9,43,18,62)
AAA_factor <- cut(x = AAA_rank, breaks = c(0, 20, 40, 60, 80, 100))
head(AAA_factor)

# Create 4 buckets for AAA_rank using cut()
AAA_factor <- cut(x = AAA_rank, breaks = c(0,25,50,75,100))

# Rename the levels 
levels(AAA_factor)<-c("low","medium","high","very_high")

# Print AAA_factor
AAA_factor

# Plot AAA_factor
plot(AAA_factor)

```
Create an ordered factor
Look at the plot created over on the right. It looks great, but look at the order of the bars! No order was specified when you created the factor, so, when R tried to plot it, it just placed the levels in alphabetical order. By now, you know that there is an order to credit ratings, and your plots should reflect that!

As a reminder, the order of credit ratings from least risky to most risky is:

AAA, AA, A, BBB, BB, B, CCC, CC, C, D
To order your factor, there are two options.

When creating a factor, specify ordered = TRUE and add unique levels in order from least to greatest:

```{r}

# Use unique() to find unique words
unique(credit_rating)

# Create an ordered factor
credit_factor_ordered <- factor(credit_rating, ordered = TRUE, levels = c("AAA","AA","BB","B","CCC"))

# Plot credit_factor_ordered
plot(credit_factor_ordered)
```

Subsetting a factor
You can subset factors in a similar way that you subset vectors. As usual, [ ] is the key! However, R has some interesting behavior when you want to remove a factor level from your analysis. For example, what if you wanted to remove the AAA bond from your portfolio?

credit_factor

[1] AAA AA  A   BBB AA  BBB A  
Levels: BBB < A < AA < AAA

credit_factor[-1]

[1] AA  A   BBB AA  BBB A  
Levels: BBB < A < AA < AAA
R removed the AAA bond at the first position, but left the AAA level behind! If you were to plot this, you would end up with the bar chart over to the right. A better plan would have been to tell R to drop the AAA level entirely. To do that, add drop = TRUE:

credit_factor[-1, drop = TRUE]

[1] AA  A   BBB AA  BBB A  
Levels: BBB < A < AA

```{r}

# Remove the A bonds at positions 3 and 7. Don't drop the A level.
keep_level <- credit_factor[-c(3,7)]

# Plot keep_level
plot(keep_level)

# Remove the A bonds at positions 3 and 7. Drop the A level.
drop_level <-credit_factor[-c(3,7),drop="TRUE"]

# Plot drop_level
plot(drop_level)

```
stringsAsFactors
Do you remember back in the data frame chapter when you used str() on your cash data frame? This was the output:

str(cash)

'data.frame':    3 obs. of  3 variables:
 $ company  : Factor w/ 2 levels "A","B": 1 1 2
 $ cash_flow: num  100 200 300
 $ year     : num  1 3 2
See how the company column has been converted to a factor? R's default behavior when creating data frames is to convert all characters into factors. This has caused countless novice R users a headache trying to figure out why their character columns are not working properly, but not you! You will be prepared!

```{r}

# Variables
credit_rating <- c("AAA", "A", "BB")
bond_owners <- c("Dan", "Tom", "Joe")

# Create the data frame of character vectors, bonds
bonds <-data.frame(credit_rating,bond_owners,stringsAsFactors = FALSE)


# Use str() on bonds
str(bonds)

# Create a factor column in bonds called credit_factor from credit_rating
bonds$credit_factor <- factor(bonds$credit_rating, ordered = TRUE, levels = c("AAA","A","BB"))

# Use str() on bonds again
str(bonds)

```

Create a list
Just like a grocery list, lists in R can be used to hold together items of different data types. Creating a list is, you guessed it, as simple as using the list() function. You could say that a list is a kind of super data type: you can store practically any piece of information in it! Create a list like so:

words <- c("I <3 R")
numbers <- c(42, 24)

my_list <- list(words, numbers)

my_list

[[1]]
[1] "I <3 R"

[[2]]
[1] 42 24
Below, you will create 

```{r}
# List components
name <- "Apple and IBM"
apple <- c(109.49, 109.90, 109.11, 109.95, 111.03)
ibm <- c(159.82, 160.02, 159.84, 160.35, 164.79)
cor_matrix <- cor(cbind(apple, ibm))

# Create a list
portfolio <- list(name,apple,ibm,cor_matrix)

# View your first list
portfolio
```
Named lists
Knowing how forgetful you are, you decide it would be important to add names to your list so you can remember what each element is describing. There are two ways to do this!

You could name the elements as you create the list with the form name = value:

my_list <- list(my_words = words, my_numbers = numbers)
Or, if the list was already created, you could use names():

my_list <- list(words, numbers)
names(my_list) <- c("my_words", "my_numbers") 
Both would result in:

my_list

$my_words
[1] "I <3 R"

$my_numbers
[1] 42 24

```{r}

# Add names to your portfolio
names(portfolio)<-c("portfolio_name", "apple", "ibm", "correlation")

# Print portfolio
portfolio
```

Access elements in a list
Subsetting a list is similar to subsetting a vector or data frame, with one extra useful operation.

To access the elements in the list, use [ ]. This will always return another list.

my_list[1]

$my_words
[1] "I <3 R"

my_list[c(1,2)]

$my_words
[1] "I <3 R"

$my_numbers
[1] 42 24
To pull out the data inside each element of your list, use [[ ]].

my_list[[1]]

[1] "I <3 R"
If your list is named, you can use the $ operator: my_list$my_words. This is the same as using [[ ]] to return the inner data.

```{r}

# Second and third elements of portfolio
portfolio[c(2,3)]

# Use $ to get the correlation data
portfolio$correlation

```
Adding to a list
Once you create a list, you aren't stuck with it forever. You can add new elements to it whenever you want! Say you want to add your friend Dan's favorite movie to your list. You can do so using $ like you did when adding new columns to data frames.

my_list$dans_movie <- "StaR Wars"

my_list

$my_words
[1] "I <3 R"

$my_numbers
[1] 42 24

$dans_movie
[1] "StaR Wars"
You could have also used c() to add another element to the list: c(my_list, dans_movie = "StaR Wars"). This can be useful if you want to add multiple elements to your list at once.

```{r}

# Add weight: 20% Apple, 80% IBM
portfolio$weight <- c(apple = 0.2, ibm = 0.8)

# Print portfolio
portfolio


# Change the weight variable: 30% Apple, 70% IBM
portfolio$weight<- c(apple = 0.3, ibm = 0.7)

# Print portfolio to see the changes
portfolio

```
Removing from a list
The natural next step is to learn how to remove elements from a list. You decide that even though Dan is your best friend, you don't want his info in your list. To remove dans_movie:

my_list$dans_movie <- NULL

my_list

$my_words
[1] "I <3 R"

$my_numbers
[1] 42 24
Using NULL is the easiest way to remove an element from your list! If your list is not named, you can also remove elements by position using my_list[1] <- NULL or my_list[[1]] <- NULL

```{r}

# Take a look at portfolio
portfolio

# Remove the microsoft stock prices from your portfolio
portfolio$microsoft<-NULL
portfolio
```

Split it
Often, you will have data for multiple groups together in one data frame. The cash data frame was an example of this back in Chapter 3. There were cash_flow and year columns for two groups (companies A and B). What if you wanted to split up this data frame into two separate data frames divided by company? In the next exercise, you will explore why you might want to do this, but first let's explore how to make this happen using the split() function.

Create a grouping to split on, and use split() to create a list of two data frames.

grouping <- cash$company
split_cash <- split(cash, grouping)

split_cash 

$A
  company cash_flow year
1       A      1000    1
2       A      4000    3
3       A       550    4

$B
  company cash_flow year
4       B      1500    1
5       B      1100    2
6       B       750    4
7       B      6000    5

```{r}

# Define grouping from year
grouping <- cash$year

# Split cash on your new grouping
split_cash <- split(cash,grouping)

# Look at your split_cash list
split_cash

# Unsplit split_cash to get the original data back.
original_cash <- unsplit(split_cash,grouping)

# Print original_cash
original_cash

```

Split-Apply-Combine
A common data science problem is to split your data frame by a grouping, apply some transformation to each group, and then recombine those pieces back into one data frame. This is such a common class of problems in R that it has been given the name split-apply-combine. In Intermediate R for Finance, you will explore a number of these problems and functions that are useful when solving them, but, for now, let's do a simple example.

Suppose, for the cash data frame, you are interested in doubling the cash_flow for company A, and tripling it for company B:

grouping <- cash$company
split_cash <- split(cash, grouping)

# We can access each list element's cash_flow column by:
split_cash$A$cash_flow
[1] 1000 4000  550

split_cash$A$cash_flow <- split_cash$A$cash_flow * 2
split_cash$B$cash_flow <- split_cash$B$cash_flow * 3

new_cash <- unsplit(split_cash, grouping)
Take a look again at how you access the cash_flow column. The first $ is to access the A element of the split_cash list. The second $ is to access the cash_flow column of the data frame in A.

```{r}

# Print split_cash
split_cash

# Print the cash_flow column of B in split_cash
split_cash$B$cash_flow

# Set the cash_flow column of company A in split_cash to 0
split_cash$A$cash_flow <-0

# Use the grouping to unsplit split_cash
cash_no_A <- unsplit(split_cash,grouping)

# Print cash_no_A
cash_no_A

```
Attributes
You have made it to the last exercise in the course! Congrats! Let's finish up with an easy one.

Attributes are a bit of extra metadata about your data structure. Some of the most common attributes are: row names and column names, dimensions, and class. You can use the attributes() function to return a list of attributes about the object you pass in. To access a specific attribute, you can use the attr() function.

Exploring the attributes of cash:

attributes(cash)

$names
[1] "company"   "cash_flow" "year"     

$row.names
[1] 1 2 3 4 5 6 7

$class
[1] "data.frame"

attr(cash, which = "names")

[1] "company"   "cash_flow" "year"     

```{r}

# my_matrix and my_factor
my_matrix <- matrix(c(1,2,3,4,5,6), nrow = 2, ncol = 3)
rownames(my_matrix) <- c("Row1", "Row2")
colnames(my_matrix) <- c("Col1", "Col2", "Col3")

my_factor <- factor(c("A", "A", "B"), ordered = T, levels = c("A", "B"))

# attributes of my_matrix
attributes(my_matrix)

# Just the dim attribute of my_matrix
attr(my_matrix,"dim")

# attributes of my_factor
attributes(my_factor)

```